A weakly structured stem for human origins in Africa

### 2023-02-26 paper.tex

\begin{table}[t!]
    \centering
    \scriptsize
    \begin{tabular}{lllrrrr}
        \toprule
        \textbf{Likelihood} & \textbf{Label} & \textbf{Population Pair} &
            \textbf{Divergence} & \textbf{Migration rate} & \textbf{Migration} \\
        & & & 
            \textbf{Time (ka)} & \textbf{per generation} & \textbf{duration (kyr)} \\
        \midrule
        \textbf{Continuous Model} & & & & & \\
        $LL= -115,300$ & a & Stem 1, Stem 2 & 1,223 & 6.26e-5 & 1,089 \\
        (Table~\ref{tab:supp-continuous-migration}) & b & Stem 2, Nama & NA & 5.85e-5 & 129 \\
        & c, d & Stem 2, Other Africans* & NA & 3.10e-5, \textbf{1.62e-4} & 129, 55 \\
        & e, f & Nama, Other Africans* & 135 & 4.1e-5, 9.2e-6 & 134, 60 \\
        & g & Mende, East Africans & 60 & \textbf{2.13e-4} & 60 \\
        & h & East Africans, British & 50 & 4.16e-5 & 50 \\
        & i & Gumuz, Amhara/Oromo & 12 &\textbf{3.37e-4} & 12 \\
        \textbf{Merger Model} & & & & & \\
        $LL= -101,600$ & a & Stem 1, Stem 2 & 1,692 & \textbf{1.26e-4} & 1,213 \\
        (Table~\ref{tab:supp-merger-with-stem-migration}) & -- & Stem 1S, Stem 1E & 479 & 0 (fixed) & -- \\
        & b & Stem 2 to Nama & 119 & \textbf{0.70} & pulse \\
        & c & Stem 2 to Stem 1E & 98 & \textbf{0.52} & pulse \\
        & d & Stem 2 to Mende & 25 & \textbf{0.19} & pulse \\
        & e, f & Nama, Other Africans* & 119 & 4.5e-5, 9.8e-6 & 120, 60 \\
        & g & Mende, East Africans* & 60 & \textbf{1.97e-4} & 60 \\
        & h & East Africans, British & 50 & 3.82e-5 & 50 \\
        & i & Gumuz, Amhara/Oromo &12 & \textbf{3.59e-4} & 12 \\
        \bottomrule
    \end{tabular}
    \caption{
        \textbf{Migration and divergence parameters from best fit models.}
        Labeled migration rates correspond to symmetric continuous migration
        bands shown in Figure~\ref{fig:best-fit-models}. Both the continuous migration and
        merger models inferred a relatively deep split of human stem branches,
        though they were connected by ongoing migration that maintained their
        genetic similarity. Bold text indicates migration rates above $10^{-4}$.
        In both models, the branch ancestral to the Nama shares a common ancestral population
        with the other African groups $\sim120-135$ka. Following this divergence,
        the population ancestral to other African groups branches into West and East African
        groups $60$ka.
        *Migration rates and durations are shown between branches ancestral to 
        1) Nama and East Africans and their ancestors, and
        2) Nama and Mende, respectively.
        ``Divergence times'' correspond to the most recent common ancestral population,
        and does not account for continuous migration or earlier reticulations.
    }
    \label{tab:migration-rates}
\end{table}

@@

\subsection*{Population structure formation in Africa}

>>>>

\begin{table}[t!]
    \centering
    \scriptsize
    \begin{tabular}{lllrrrr}
        \toprule
        \textbf{Likelihood} & \textbf{Label} & \textbf{Population Pair} &
            \textbf{Divergence} & \textbf{Migration rate} & \textbf{Migration} \\
        & & & 
            \textbf{Time (ka)} & \textbf{per generation} & \textbf{duration (kyr)} \\
        \midrule
        \textbf{Continuous Model} & & & & & \\
        $LL= -115,300$ & a & Stem 1, Stem 2 & 1,223 & 6.26e-5 & 1,089 \\
        (Table~\ref{tab:supp-continuous-migration}) & b & Stem 2, Nama & NA & 5.85e-5 & 129 \\
        & c, d & Stem 2, Other Africans* & NA & 3.10e-5, \textbf{1.62e-4} & 129, 55 \\
        & e, f & Nama, Other Africans* & 135 & 4.1e-5, 9.2e-6 & 134, 60 \\
        & g & Mende, East Africans & 60 & \textbf{2.13e-4} & 60 \\
        & h & East Africans, British & 50 & 4.16e-5 & 50 \\
        & i & Gumuz, Amhara/Oromo & 12 &\textbf{3.37e-4} & 12 \\
        \textbf{Merger Model} & & & & & \\
        $LL= -101,600$ & a & Stem 1, Stem 2 & 1,692 & \textbf{1.26e-4} & 1,213 \\
        (Table~\ref{tab:supp-merger-with-stem-migration}) & -- & Stem 1S, Stem 1E & 479 & 0 (fixed) & -- \\
        & b & Stem 2 to Nama & 119 & \textbf{0.70} & pulse \\
        & c & Stem 2 to Stem 1E & 98 & \textbf{0.52} & pulse \\
        & d & Stem 2 to Mende & 25 & \textbf{0.19} & pulse \\
        & e, f & Nama, Other Africans* & 119 & 4.5e-5, 9.8e-6 & 120, 60 \\
        & g & Mende, East Africans* & 60 & \textbf{1.97e-4} & 60 \\
        & h & East Africans, British & 50 & 3.82e-5 & 50 \\
        & i & Gumuz, Amhara/Oromo &12 & \textbf{3.59e-4} & 12 \\
        \bottomrule
    \end{tabular}
    \caption{
        \textbf{Migration and divergence parameters from best fit models.}
        Labeled migration rates correspond to symmetric continuous migration
        bands shown in Figure~\ref{fig:best-fit-models}. Both the continuous migration and
        merger models inferred a relatively deep split of human stem branches,
        though they were connected by ongoing migration that maintained their
        genetic similarity. Bold text indicates migration rates above $10^{-4}$.
        In both models, the branch ancestral to the Nama shares a common ancestral population
        with the other African groups $\sim120-135$ka. Following this divergence,
        the population ancestral to other African groups branches into West and East African
        groups $60$ka.
        *Migration rates and durations are shown between branches ancestral to 
        1) Nama and East Africans and their ancestors, and
        2) Nama and Mende, respectively.
        ``Divergence times'' correspond to the most recent common ancestral population
        and does not account for continuous migration or earlier reticulations.
    }
    \label{tab:migration-rates}
\end{table}

@@

\subsection*{Formation of Population Structure in Africa}
